SubGenome Analysis

### 2017-07-27 R/SubGenome Analysis.Rmd
Mutate all useful values and summaries upfront

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(readxl)
library(plyr)#
library(dplyr)
library(tidyr)
library(reshape2)#
library(ggplot2)
```
```{r preamble, echo=FALSE}
paste("This analysis is using the file", params$inputDataFile)
```

@@

## Preprocessing to generate a tidy dataset
We parse the SynMap output file with ks/kn values (i.e. a DAGChainer file) to a flat format using a perl script (parseKsKnFile.pl) and import the data into R.  Some ks and kn values are empty, so we clean the dataset by removing all incomplete observations.

> manually perform the parse and run the import lines until pipeline is debugged
> perl parseKsKnFile.pl ../Data/6807_8082.CDS-CDS.lastz.tdd10.cs0.filtered.dag.all.go_D20_g10_A5.aligncoords.gcoords.ks > ../Data/sorghum_v1_vs_maize_v1.tab

```{r importSynMapData, include=FALSE}
cmd <- paste("perl", params$fileParser, params$synMapOutputFile)
system(cmd)
syntelogs.raw <- read_delim(params$inputDataFile, "\t", escape_double = FALSE, trim_ws = TRUE)
syntelogs.complete <- syntelogs.raw[complete.cases(syntelogs.raw),]
```
```{r reportSynMapData, echo=FALSE, results="asis"}
# Report results of import
cat(sprintf("We have imported %d lines from the parsed SynMap file and removed %d incomplete rows resulting in %d syntenic genes.", nrow(syntelogs.raw), nrow(syntelogs.raw)- nrow(syntelogs.complete), nrow(syntelogs.complete)))
```

@@

# Reconstructed ancestral chromosomes
homeologs %>% group_by(org_chr1, org_chr2) %>% summarize(genes=sum(count)) %>% spread(org_chr2, genes) -> homeologs.spread
homeologs.spread$total <- rowSums(homeologs.spread[-1], na.rm = TRUE)
syntelogs.complete %>% select(org_chr1) %>% distinct() %>% left_join(homeologs.spread) %>% replace(is.na(.), 0) %>% arrange(desc(org_chr1)) -> homeologs.totals
```
```{r reportFindSubgenomes, echo=FALSE, results="asis"}
homeologs.totals %>% knitr::kable(caption = "Table")
homeologs.totals %>% filter(total==0) %>% nrow() %>% replace(is.na(.), 0) %>% sprintf("There are %i unaligned sorghum chromosome(s) in the above table.", .) %>% cat()

```

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(readxl)
library(plyr)#
library(dplyr)
library(tidyr)
#library(reshape2)
library(ggplot2)
```
```{r preamble, echo=FALSE}
paste("This analysis is using the file", params$inputDataFile)
```

@@

## Import Data and Preprocessing
We parse the SynMap output file with ks/kn values (i.e. a DAGChainer file) to a flat format using a perl script (parseKsKnFile.pl) and import the data into R. Some ks and kn values are empty, which is to be expected.

```{r importSynMapData, include=FALSE}
syntelogs.raw <- read_delim(params$inputDataFile, "\t", escape_double = FALSE, trim_ws = TRUE)
#syntelogs.complete <- syntelogs.raw[complete.cases(syntelogs.raw),]

syntelogs.raw %>% 
  group_by(block, org_chr1, org_chr2) %>% 
  summarise(median_ks=median(ks, na.rm=TRUE), geneCount=n()) %>% 
  left_join(syntelogs.raw, ., by = c("block", "org_chr1", "org_chr2")) %>% 
  mutate(gene_length1=abs(stop1-start1)) %>% 
  mutate(gene_length2=abs(stop2-start2)) %>% 
  mutate(chr1=regmatches(org_chr1, regexpr("\\d*$",org_chr1))) %>% 
  mutate(chr2=regmatches(org_chr2, regexpr("\\d*$",org_chr2))) %>% 
  mutate(id=paste(block, org_chr1, org_chr2, sep="||")) -> syntelogs.mutated
```
```{r reportSynMapData, echo=FALSE, results="asis"}
# Report results of import
cat(sprintf("We have imported %d lines from the parsed SynMap file and removed %d incomplete rows resulting in %d syntenic genes.", nrow(syntelogs.raw), nrow(syntelogs.raw)- nrow(syntelogs.complete), nrow(syntelogs.complete)))
```

@@

# Reconstructed ancestral chromosomes
homeologs %>% group_by(org_chr1, org_chr2) %>% summarize(genes=sum(count)) %>% spread(org_chr2, genes) -> homeologs.spread
homeologs.spread$total <- rowSums(homeologs.spread[-1], na.rm = TRUE)
syntelogs.complete %>% select(org_chr1) %>% distinct() %>% left_join(homeologs.spread) %>% replace(is.na(.), 0) %>% arrange(desc(org_chr1)) -> homeologs.totals
```
```{r reportFindSubgenomes, echo=FALSE, results="asis"}
homeologs.totals %>% knitr::kable(caption = "Table")
homeologs.totals %>% filter(total==0) %>% nrow() %>% replace(is.na(.), 0) %>% sprintf("There are %i unaligned sorghum chromosome(s) in the above table.", .) %>% cat()
```
